

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
We will apply pca on wine dataset
```{r}
wine = read.csv("wine.csv")
```
Applying PCA on relevant predictors
```{r}
pca<-prcomp(wine[,3:7],scale=TRUE)
```
Analyzing components of the output
```{r}
#Std Dev
pca$sdev
# Loadings
pca$rotation
# Principal Components
pca$x
pca$center
pca$scale
```
Creating biplot
```{r}
biplot(pca,scale=0)
```
Calculating proportion of variance
```{r}
pr.var<-pca$sdev^2
pve<-pr.var/sum(pr.var)
pve
```
Creating scree plot and cumulative plots
```{r}
plot(pve, xlab ="Principal Component", 
     ylab ="Proportion of Variance Explained", ylim=c(0 ,1) ,type="b")
plot(cumsum (pve), xlab ="Principal Component", 
     ylab =" Cumulative Proportion of Variance Explained ", ylim=c(0 ,1), type="b")
```
  
  Building model using PC1 to PC4
```{r}
predictor<-pca$x[,1:5]
wine<-cbind(wine,predictor)
model<-lm(Price~PC1+PC2+PC3+PC4,data=wine)
summary(model)
```
#### Making Predictions  
We cannot convert test data into principal components, by applying pca. Instead we have to apply same transformations on test data as we did for train data
```{r}
wineTest = read.csv("wine_test.csv")
wineTest
wineTest[1,3:7]-pca$center
wineTest[2,3:7]-pca$center
pca_test<-predict(pca,wineTest[,3:7])
class(pca_test)
pca_test
# Converting to data frame
pca_test<-as.data.frame(pca_test)
pca_test
```
Making predictions
```{r}
pred_pca<-predict(object = model, newdata=pca_test)
pred_pca
wineTest$Price
```
### Calculating out of sample `R^2`
```{r}
SSE<- sum((pred_pca-wineTest$Price)^2)
SST<-sum((wineTest$Price-mean(wine$Price))^2)
Rsq<-1-SSE/SST
```
